---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 1: Getting sttarted'
author: "Dr. Nico Gradwohl"
date: "`r Sys.Date()`"
output: html_document
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for two.

1. Upload your assignments to the joint git repository. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


# Task 1: Getting set and ready with R and RStudio 

This task 

*Note*: 
You may also work in python and use a Jupyter notebook, howerver, I can help you less with that.

```{r instructions-install-packages, eval=FALSE}
install.packages("dplyr")

install.packages("ggplot2")
```


# Task 2: A simple data model 

The following code-chunk simulates some data (for now you do not need to understand how) and outputs a histogram.

```{r simulate-data}

```

Find any model for this data. **TODO!**

```{r task1}

```


**TODO: BASE ON MCELREATH EXAMPLE?**


# Bonus task 
