---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 1: Getting sttarted'
author: "Dr. Nico Gradwohl"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for two.

1. Upload your assignments to the joint git repository. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


# Task 1: Getting set and ready with R and RStudio 

This task 

*Note*: 
You may also work in python and use a Jupyter notebook, howerver, I can help you less with that.

```{r instructions-install-packages, eval=FALSE}
install.packages("dplyr")

install.packages("ggplot2")
```


# Task 2: A simple data model 

The following code-chunk simulates some data (for now you do not need to understand how) and outputs a histogram.

```{r TODO-simulate-data}
# Taken from McElreath (YYYY)
    # 5M1: Spurious correlation:
        set.seed(2022)
        n <- 100
        temp <- rnorm(n)
        shark <- rnorm(n, temp)  # deviations around the temperature means.
        ice_cream <- rnorm(n, -shark)
        
        plot(shark, temp)
        plot(ice_cream, temp)
            

            
    # 5M2: Masked relationship:
        n <- 100
        learn_speed <- rnorm(n)
        # TODO: Make one relationship non-linear?
        alcohol <- rnorm(n, learn_speed)  # assuming that more intelligent people drink more
        intelligence <- rnorm(n, alcohol - learn_speed)
        
        plot(intelligence, learn_speed)
        plot(alcohol, learn_speed)
            

```

Find any model for this data. **TODO!**

* Model for temperature or learn speed
    + plot the data for your model 
    + fit a model 


## Answer Task 2:

```{r task2}
            lm(ice_cream ~ temp)
            lm(ice_cream ~ shark)
            lm(ice_cream ~ temp + shark)
            
            lm(learn_speed ~ alcohol)
            lm(learn_speed ~ intelligence)
            lm(learn_speed ~ alcohol + intelligence)
```


**TODO: BASE ON MCELREATH EXAMPLE?**


# Bonus task 

**Fitting a non-linear model?**
**Detecting non-linear relationship?**

* Name (potentially hidden) assumptions 
