---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 1: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for two.

1. Upload your assignments to the joint git repository. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


# Task 1: Getting set and ready with R and RStudio 

This task 

*Note*: 
You may also work in python and use a Jupyter notebook, however, I can help you less with that.

```{r instructions-install-packages, eval=FALSE}
install.packages("dplyr")

install.packages("ggplot2")
```


# Task 2: A simple data model 

The following code-chunk simulates some data (for now you do not need to understand how -- but try to understand the code anyway).
They describe a fictitious relationship between alcohol consumption (variable `alcohol`) and learn speed (variable `learn_speed`, i.e., how quickly someone grasps a concept).


```{r TODO-simulate-data}
# Taken from McElreath (YYYY)
    # Simulate data:
        set.seed(42)

        n <- 100
        learn_speed <- rnorm(n)
        alcohol <- rnorm(n, learn_speed)  # assuming that more intelligent people drink more
```



### Task 2a: Plot the relationship between alcohol and learning speed (`learn_speed`)  

```{r task2a}
# ENTER YOUR SOLUTIONS TO TASK 2a HERE 
```

### Task 2b: Find a model for the relationship between alcohol and learning speed  


```{r task2b}
# ENTER YOUR SOLUTION TO TASK 2b HERE 
```

### Task 2c: What do the model parameters mean? 

*[WRITE YOUR SOLUTION TO TASK 2c HERE]*


### Task 2d (optional): 

Now, we also got some data for individuals' `intelligence`.

Look at the relationship between alcohol, intelligence and loearning speed). 
What might be happening here?

```{r bonus-task2d}
  intelligence <- rnorm(n, alcohol - learn_speed)
```

*[WRITE YOUR SOLUTION TO THE BONUS TASK HERE]*



