---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 1: Getting started'
author: "Dr. Nico Gradwohl"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for two.

1. Upload your assignments to the joint git repository. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


# Task 1: Getting set and ready with R and RStudio 

This task 

*Note*: 
You may also work in python and use a Jupyter notebook, howerver, I can help you less with that.

```{r instructions-install-packages, eval=FALSE}
install.packages("dplyr")

install.packages("ggplot2")
```


# Task 2: A simple data model 

The following code-chunk simulates some data (for now you do not need to understand how) and outputs a histogram.
They describe a ficticious relationship between alcohol and intelligence.

```{r TODO-simulate-data}
# Taken from McElreath (YYYY)

      set.seed(2022)
      n <- 100
      temp <- rnorm(n)
      shark <- rnorm(n, temp)  # deviations around the temperature means.
      ice_cream <- rnorm(n, -shark)
      
      plot(shark, ice_cream)
      plot(shark, temp)
      plot(ice_cream, temp)
      
      
    # 5M2: Masked relationship:
        n <- 100
        learn_speed <- rnorm(n)
        # TODO: Make one relationship non-linear?
        alcohol <- rnorm(n, learn_speed)  # assuming that more intelligent people drink more
        intelligence <- rnorm(n, alcohol - learn_speed)
        
        plot(intelligence, learn_speed)
        plot(alcohol, learn_speed)
            

```

Find any model for this data. **TODO!**

* Model for temperature or learn speed
    + plot the data for your model 
    + fit a model 


### Task 2a: Plot the relationship between alcohol and learning speed (`learn_speed`)  

```{r task2a}
# ENTER YOUR SOLUTIONS TO TASK 2a HERE 
```

### Task 2b: Find a model for the relationship between alcohol and learning speed  


```{r task2b}
# ENTER YOUR SOLUTION TO TASK 2b HERE 
```

### Task 2c: What do the model parameters mean? 

*[WRITE YOUR SOLUTION TO TASK 2c HERE]*


### Task 2d (optional): Now look at the relationship between alcohol and intelligence. 

```{r bonus-task2d}

```

*[WRITE YOUR SOLUTION TO THE BONUS TASK HERE]*



