---
title: 'Computational modelling of judgments, decisions and cognition -- Materials
author: "Dr. Nico Gradwohl"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.

thlg <- theme(axis.title = element_text(size = 16),
              axis.text = element_text(size = 14),
              legend.title = element_text(size = 14),
              legend.text = element_text(size = 12))
```


# Ressources for data etc. 

* Exercises for book. https://psy-farrell.github.io/computational-modelling//exercises/ 

* Code from book: https://osf.io/ejfm4/files/osfstorage

* Data and code from Tump et al. (2020) https://osf.io/ejfm4/files/osfstorage



# Session 0 

## Example: Task performance 

```{r, fig-forget}

# Forgetting data:
  N <- 150
  t <- 0:N

  # Power:
    Ap <- 1000
    Bp <- 500
    bet <- 0.4
  
    RTp <- Ap + Bp * (t + 1)^(-bet) + rnorm(N + 1, 0, 50)

  # Exponential:
    Ae <- 1000
    Be <- 500
    alp <- 0.6
  
    RTe <- Ae + Be * exp(-alp * t) + rnorm(N + 1, 0, 50)

  # Linear:

# Plotting:
  pow_lin_dat <- data.frame(t = t, RTp = RTp, RTe = RTe,
              pred_p = Ap + Bp * (t + 1)^(-bet),
              pred_e = Ae + Be * exp(-alp * t)) %>%
    ggplot(aes(x = t)) +
    geom_point(aes(y = RTp), size = 3, alpha = 0.5) +
    labs(y = "RT (ms)", linetype = "") +
    geom_line(aes(y = pred_p, linetype = "Power law"), linewidth = 1, color = NA) +
    # geom_line(aes(y = pred_e, linetype = "Exponetial law"), linewidth = 1) +
    # Linear:
    geom_smooth(aes(y = RTp, linetype = "Linear"), method = "lm", se = FALSE, linewidth = 1, color = NA) +
    theme_bw() +
    thlg

  pow_lin_dat

  ggsave("_figures/pow_lin_dat.png", height = 12, width = 16, unit = "cm")


  pow_lin <- pow_lin_dat +
    # geom_line(aes(y = pred_e, linetype = "Exponetial law"), linewidth = 1) +
    # Linear:
    geom_smooth(aes(y = RTp, linetype = "Linear"), method = "lm", se = FALSE, linewidth = 1, color = "black")

  
  pow_lin
  ggsave("_figures/lin.png", height = 12, width = 16, unit = "cm")


  pow_lin +
    geom_line(aes(y = pred_p, linetype = "Power law"), linewidth = 1)

  ggsave("_figures/pow_lin.png", height = 12, width = 16, unit = "cm")

```





## Tax cut 

```{r}
# Average tax cut of $ 1,083 
# 
# But: 80% receive less than this cut, half (around 45 Million people) receive less than $ 100, $ 30,127 for the top 1%

# Try with lognormal:
  curmu <- 0.01
  cursd <- 10000
  
  x <- seq(0, curmu + cursd, by = 0.1)
  plot(x, dlnorm(x, log(curmu), log(cursd)), type = "l")
  
  
  1- plnorm(c(30127), log(curmu), log(cursd))
  plnorm(c(1083, 100), log(curmu), log(cursd))
  
  # Target probabilities:
  c(0.01, 0.8, 0.5)
  
  # Use a pareto distribution:
    # Equations:
    # x_m = 1/(1 - 0.5)^(-1/a) = 100
    # x_m = 1/(1 - 0.8)^(-1/a) = 1083
    # x_m = 1/(1 - 0.01)^(-1/a) = 30127
    # a <- 1083/(1083-x_m)


    x_m <- 200 

    a <- 1083/(1083-x_m)
    
    
    x <- seq(1, 10000, by = 0.2)

    pout <- (a * x_m^a)/(x^(a+1))  # PDF.

    (mpar <- (a * x_m)/(a - 1))
    # To have a mean of 1083 a can be: a = mpar/(mpar - x_m)
    
    (qtpar08 <- x_m * (1 - 0.8)^(-1/a))  # 30127

    (qtpar05 <- x_m * (1 - 0.5)^(-1/a))
    

    data.frame(x = x, y = pout) %>%
      ggplot(aes(x = x, y = y)) +
      geom_line(linewidth = 1) +
      geom_vline(aes(linetype = "mean"), xintercept = mpar) +
      geom_vline(aes(linetype = "mean"), xintercept = mpar) +
      scale_x_continuous(trans = "log10") +
      thlg +
      theme_bw()
  

```



## Linear model 

```{r lin-rel}
# Simple linear relationship:
  n <- 100
  set.seed(09234)
  learn_speed <- rnorm(n)
  alcohol <- rnorm(n, learn_speed)  # assuming that more intelligent people drink more  
```


```{r lin-dat}

  plt_alc_learn <- data.frame(alcohol = alcohol, learn_speed = learn_speed) %>%
    ggplot(aes(x = alcohol, y = learn_speed)) +
    geom_point(size = 3) +
    labs(x = "Alcohol Consumption", y = "Learn Speed", linetype = "") +
    theme_bw() +
    thlg

# Show and save:
plt_alc_learn

ggsave("_figures/alc_learn_raw.png", height = 12, width = 12, unit = "cm")
```


```{r mean-mod}
plt_alc_learn +
    geom_hline(yintercept = mean(learn_speed), linewidth = 1)

ggsave("_figures/alc_learn_mean.png", height = 12, width = 12, unit = "cm")
```


```{r lin-mod-simple}
  plt_alc_learn +
      # Linear:
      geom_abline(slope = 1, linewidth = 1) +
      annotate(geom = "polygon", x = c(1, 0, 1), y = c(1, 0, 0),
              linetype = "dashed", color = "black",
              alpha = 0.3)
  
  ggsave("_figures/alc_learn_lm.png", height = 12, width = 12, unit = "cm")


# Fitted model:
  fit <- lm(learn_speed ~ alcohol)
  summary(fit)

  ic <- coef(fit)["(Intercept)"]   # intercept.
  slp <- coef(fit)["alcohol"]

plt_alc_learn +
    # Linear:
    geom_smooth(method = "lm", se = FALSE, linewidth = 1, color = "black") +
    geom_abline(intercept = ic, slope = slp, linewidth = 1) +
    annotate(geom = "polygon", x = c(1, 0, 1), y = c(1, 0, 0) * slp + ic,
            linetype = "dashed", color = "black",
            alpha = 0.3) +
    annotate(geom = "text", x = 1, y = slp/2 + ic, label = round(slp, 2), size = 6, hjust = 0)

ggsave("_figures/alc_learn_fit.png", height = 12, width = 12, unit = "cm")
```



# Session 1 

## Mass vs. Density 

```{r mass-dens}
n_correct <- 0:10
pvals <- dbinom(n_correct, size = 10, prob = .75)
pmass <- data.frame(n_correct = factor(n_correct), pn = pvals) %>%
  ggplot(aes(x = n_correct, y = pn)) +
  geom_col() +
  # scale_x_binned(n.breaks = 11, show.limits = TRUE, right = FALSE) +
  theme_classic()

pmass
ggsave("_figures/pmass.png", height = 12, width = 12, unit = "cm")

pmass +
  annotate(geom = "segment", 
           x = 8, y = 0, yend = pvals[n_correct == 7], linetype = "dashed", linewidth = 1) +
  annotate(geom = "segment", 
           x = 0, xend = 8, y = pvals[n_correct == 7], linetype = "dashed", linewidth = 1) +
  annotate(geom = "point", x = 8, y = pvals[n_correct == 7], size = 3)

ggsave("_figures/pmass_lik.png", height = 12, width = 12, unit = "cm")

# Density:
tr <- seq(0, 30, by = 0.01)
dvals <- dlnorm(tr, 2, 1.7)
pdens <- data.frame(t = tr, dens = dvals) %>%
  ggplot(aes(x = tr, y = dens)) +
  geom_line(linewidth = 1) +
  theme_classic()

pdens
ggsave("_figures/pdens.png", height = 12, width = 12, unit = "cm")

pdens +
  annotate(geom = "segment", 
           x = 7, y = 0, yend = dvals[tr == 7], linetype = "dashed", linewidth = 1) +
  annotate(geom = "segment", 
           x = 0, xend = 7, y = dvals[tr == 7], linetype = "dashed", linewidth = 1) +
  annotate(geom = "point", x = 7, y = dvals[tr == 7], size = 3)
ggsave("_figures/pdens_lik.png", height = 12, width = 12, unit = "cm")
```



## Visualize Prospect Theory 

Value function 

```{r valfun}
# Define function:
value_function <- function(x, alp, bet = alp, lamb){
  ifelse(x >= 0, x^alp, -lamb * (-x)^bet)
}


# Visualize:
  x <- seq(-100, 100, by = 1)  # define values.


# Create combinations of different parameters:
  dm <- expand.grid(alp = c(0.1, 0.5, 0.8), lambda = c(1, 2))

# Get data for each combination:
  outlist <- apply(dm, MARGIN = 1, FUN = function(rw){
    data.frame(x = x, V = value_function(x, alp = rw[1],
      lamb = rw[2]), alp = rw[1], lamb = rw[2]) 
  })

  outdat <- do.call(rbind, outlist)  # combine the list into a single data.frame.

outdat %>%
  ggplot(aes(x = x, y = V, color = alp, linetype = factor(lamb), group = interaction(lamb, alp))) +
  geom_line(linewidth = 1) +
  scale_color_distiller(palette = "Spectral") +
  theme_classic()

ggsave("_figures/valfun.png", height = 12, width = 12, unit = "cm")
```


## Visualize MAB 

```{r generate-pdat}
probdat <- seq(0.05, 0.95, by = 0.15)  # probabilities.

# FUnction to generate rewards:
curprob <- probdat[1]  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.

# Simple Q-learning:-----------------
  get_cmat <- function(choice, input){
    
    # Input matrix with row and column indices and values:
    inp <- as.matrix(cbind(1:length(choice), choice, input))
    
    # Get own rewards for each choice:
    cmat <- matrix(0, ncol = 2, nrow = length(choice)) 
    cmat[inp[,1:2]] <- inp[, 3]
    
    return(cmat)
    
  }

  # Utility function:
  utpay <- function(rew, choice) {
  
    # Input matrix with row and column indices and values:
    paymat <- get_cmat(choice = choice, input = rew)
    
    return(paymat)

  }

update_q <- function(qprev,  # vector of previous q values
                     ut,  # vector of rewards / utilities.
                     alp,  # alpha
                     update_unchosen = TRUE,
                     choices = NA
)  {
  
  
  pred_err <- ut - qprev  # prediction error.
  
  # Should the unchosen option be updated?
  if(!update_unchosen){
    # Set prediction error == 0 for unchosen option:
    pred_err[3 - choices] <- 0
      # Note: Replace 2 by length(options) for more than 2 options.
      # Alternatively: Update the unchosen and rewarded option with 1? 
    # ut[cbind(1:3, choices + 1)]  # rewards from chosen.
    
  }
  
  qnew <- qprev + alp * pred_err
  
  return(qnew)
  
}


  simfun <- function(theta, competition = TRUE, op_prob = c(0.5, 0.5),
                       utfun = utpay, nind = 3, maxtr = 200, maxrew = 1){
      
      ## 4.1 Get parameters:  
      lpar <- theta[1]  # get corresponding learning parameter phi/alpha.
      soft <- theta[length(theta)]  # softmax parameter is always last in list!.
      
      # For mimicking Schulze et al., the softmax paramter is transformed
      # (we use soft instead of c to avoid collision):
      
      soft <- 3^(10 * soft) - 1  
      
      # First round: -------------------------------------------------------
      # Choices (rows are rounds, columns individuals):
      cmat <- matrix(sample(c(0, 1), size = nind, prob = c(0.5, 0.5), replace = TRUE),
                     nrow = 1)
      
      # Rewards (rows are rounds, columns individuals):
      rew_choice <- sample(c(0, 1), size = 1,
                           prob = op_prob)
      rmat <- ifelse(rew_choice == cmat & any(rew_choice == cmat),  
                      # if any chose the rewarded option for all.
                     competition * maxrew/sum(rew_choice == cmat) +
                       (1 - competition) * maxrew,  # maximum reward distributed.  
                     0)
      
      # utfun(rew = rmat[1, ], choice = cmat[1,])
      utmat <- array(rep(0, 2 * nind * maxtr), dim = c(nind, 2, maxtr))  # utility matrix in given round.
      # Individuals are rows, opotions are columns.
      
      utmat[, rew_choice + 1, 1] <- rmat  # add rewarded option.
      # TODO: Translate utuitlity if necessary.
      utmat[,,1:2]
      
      
      ## 3.3 Initialize option value matrix:
      optval_mat <- array(rep(NA, 2 * nind * maxtr), dim = c(nind, 2, maxtr))   # option values for each choice
      optval_mat[,,1] <- 0
      
      
      
    # 4.2 Loop over subsequent trials: ---------------------------------
      for(tr in 2:maxtr) {
    
    
        # Learning:
        ## 4.2.1 Get data from previous trial:
        choices <- cmat[tr - 1,]  # choices in previous round.
        pi_cur <- utmat[, , tr - 1]  # get utilities from previous trial.
    
        ## 4.2.2 Update learned values:
    
        # Which formulation is used?
        # TODO: Probably this is actually obsolete as both functions are essentially the same!
    
        # a) A learning rule:
        # Matrix of option values in current round:
        optval_mat_tr <- update_q(qprev = optval_mat[,,tr - 1],
                                  ut = pi_cur,
                                  alp = lpar)
    
        if(is.null(dim(optval_mat_tr))){
          dim(optval_mat_tr) <- c(nind, 2)
        }
    
        # Save option value matrix:
        optval_mat[,,tr] <- optval_mat_tr  # get a matrix of q-values.
    
        # Choices:
        # subtraction of each row's maximum to avoid overflow problems.
        optval_mat_tr <- t(apply(optval_mat_tr, MAR = 1, function(x) x - max(x)))
    
        # Adapted from Charley:
        p <- exp(optval_mat_tr * soft)  # Softmax rule.
        # TODO: Use division or multiplication?
    
        p <- p/rowSums(p)  # get probabilities for each option.
    
        # Get a choice for each individual:
        # sample(matrix(rep(c(0, 1), each = nind), ncol = 2), size = nind, replace = TRUE, prob = p)
    
        cvec <- apply(p, MARGIN = 1, FUN = function(p){
          sample(c(0, 1), size = 1, prob = p)
        })
    
        cmat <- rbind(cmat, cvec)
    
        # Rewards:
        # Rewards (rows are rounds, columns individuals):
        rew_choice <- sample(c(0, 1), size = 1, prob = op_prob)
        rmat <- ifelse(rew_choice == cvec & any(rew_choice == cvec),  # if any chose the rewarded option for all.
                       competition * maxrew/sum(rew_choice == cvec) +
                         (1 - competition) * maxrew/2,  
                          # maximum reward distributed (divide by 2 for non-competition).
                       0)
    
    
        utmat[, rew_choice + 1, tr] <- rmat  # add rewarded option.
        # TODO: Translate utuitlity if necessary.
    
      }  # eof. loop.
    
      return(list(cmat = cmat, optval = optval_mat))
    
    }
```

```{r run-sims}
theta_cur <- 0.1

dm <- expand.grid(theta_cur = c(0.2, 0.8),
                  p = c(0.2, 0.5, 0.9),
                  alp = 0.9)

outlist <- lapply(1:nrow(dm), FUN = function(rw){
  simout <- simfun(theta = c(dm$alp[rw], dm$theta_cur[rw]), nind = 1, op_prob = c(dm$p[rw], 1-dm$p[rw]))

  soft <- 3^(10 * dm$theta_cur[rw]) - 1  

  optval_mat_tr <- t(apply(simout$optval[1,,], MAR = 2, function(x) x - max(x)))
  oprobs <- exp(optval_mat_tr * soft)
  data.frame(tr = 1:length(simout$cmat), p = dm$p[rw], theta = dm$theta_cur[rw],
        oval = simout$optval[,2,], oprob = (oprobs/rowSums(oprobs))[,2], 
        choice = simout$cmat)
})


do.call(rbind, outlist) %>%
  ggplot(aes(x = tr, y = oprob, color = factor(p))) +
  facet_grid(theta ~ p, labeller = label_both) +
  geom_line(linewidth = 0.5, alpha = 0.7) +
  geom_point(aes(y = choice), position = position_jitter(height = 0.01)) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("_figures/rl_example.png", height = 16, width = 24, unit = "cm")
```


## Illustrate Grid search 

### From Farell & Lewandowsky 


```{r regdata}
#define parameters to generate data
nDataPts  <- 20
rho       <- .8  # correlation.
intercept <- .0

# generate synthetic data
dat <- matrix(0, nDataPts, 2)
dat[ ,2] <- rnorm(nDataPts)   
dat[ ,1] <- rnorm(nDataPts)*sqrt(1.0-rho^2) + dat[ ,2]*rho + intercept  

dat

plot(dat)
```

```{r fitreg}
lm(dat[,1] ~ dat[,2])
```

Prediction function:

```{r predreg}
# plot data and current predictions          
getregpred <- function(parms, dat) {
  regpred <- parms["b0"] + parms["b1"] * dat[ ,2]           
  
  #wait with drawing a graph until key is pressed
  par(ask=TRUE)
  plot   (dat[ ,2], type="n", las=1, ylim=c(-2,2), xlim=c(-2,2), xlab="X", ylab="Y")
  par(ask=FALSE)
  points (dat[ ,2], dat[ ,1], pch=21, bg="gray")
  lines  (dat[ ,2], regpred, lty="solid")
  
  return(regpred)
}  
```


Discrepancy function: 

```{r reg-discr}
#obtain current predictions and compute discrepancy
rmsd <- function(parms, data1) {          
  preds <- getregpred(parms, data1)        
  rmsd <- sqrt(sum((preds-data1[ ,1])^2)/length(preds))
}
```

```{r fitreg-grid}
# Define the grid:
dm <- expand.grid(b0 = seq(-0.2, 0.2, by = 0.2), b1 = seq(-1, 1, by = 0.5))

cur_discr <- 10000
bestrow <- NA

for(i in 1:nrow(dm)){
  cur_rmsd <- rmsd(parms = unlist(dm[i,]), data1 = dat)

  if(cur_rmsd < cur_discr){
    cur_discr <- cur_rmsd
    bestrow <- i
  }
}

cur_discr
dm[bestrow,]
```


```{r fitreg-nm}
# Nelder-Mead:
startParms <- c(-1., .2) 
names(startParms) <- c("b1", "b0")
#obtain parameter estimates
xout <- optim(startParms, rmsd, data1 = dat)

# Show paramters:
xout
```


### Own example

```{r}
# Choices for first parameters.
cur_choices <- outlist[[1]]$choice

mean(outlist[[1]]$choice)

# Fixed choice probability:

ps <- seq(0, 1, by = 0.1)

lls <- sapply(ps, FUN = function(p){
  ll <- dbinom(cur_choices, size = 1, p, log = TRUE)
})

deviance <- -2 * colSums(lls)
min(deviance)

ps[which.min(deviance)]
```


```{r grid-search}
dm <- expand.grid(alp = seq(0, 1, by = 0.1),
                  theta = seq(0, 1, by = 0.1))

# Choices for first parameters.
outlist[[1]]$choice
```


### Parameter variability 

Check parameter variability with bootstrapping.Uses Carpenter et al. (2008).

Experiment: 

* Participants studied 60 obscure facts (e.g., "greyhounds have the best eyesight of any dog"). 

* testing after 5 minutes, 1, 2, 7, 14, 42 days 

* Characterized by a power function

\[
p = a(b t + 1)^{-c}
\]

```{r}
#discrepancy for power forgetting function 
powdiscrep <- function (parms,rec,ri) {                    
  if (any(parms<0)||any(parms>1)) return(1e6)
  pow_pred <- parms["a"] *(parms["b"]*ri + 1)^(-parms["c"])
  return(sqrt( sum((pow_pred-rec)^2)/length(ri) ))
}                                                          

#Carpenter et al. (2008) Experiment 1
rec <- c(.93,.88,.86,.66,.47,.34) # recall proportions.
ri  <- c(.0035, 1, 2, 7, 14, 42)  # retention intervals in days.

# initialize starting values 
  sparms <-c(1,.05,.7)
  names(sparms) <- c("a","b","c")
#obtain best-fitting estimates
  pout <- optim(sparms, powdiscrep, rec=rec, ri=ri)   # Simplex
  pow_pred <- pout$par["a"] *(pout$par["b"]*c(0:max(ri)) + 1)^(-pout$par["c"])  

#plot data and best-fitting predictions
  x11()
  par(cex.axis=1.2,cex.lab=1.4)
  par(mar=(c(5, 5, 3, 2) + 0.1),las=1)
  plot(ri,rec, 
       xlab = "Retention Interval (Days)", 
       ylab = "Proportion Items Retained",  
       ylim=c(0.3,1),xlim=c(0,43),xaxt="n",type="n")
  lines(c(0:max(ri)),pow_pred,lwd=2)
  points(ri,rec,pch=21, bg="dark grey",cex=2)
  dev <- pow_pred[ri+1]
  for (x in c(1:length(ri))) {
    lines(c(ri[x],ri[x]),c(dev[x],rec[x]),lwd=1)
    }
  axis(1,at=c(0:43))
```


```{r bootstrap}
# perform bootstrapping analysis
ns  <- 55
nbs <- 1000
bsparms <- matrix(NA,nbs,length(sparms))

# Predictions:
bspow_pred <- pout$par["a"] *(pout$par["b"]*ri + 1)^(-pout$par["c"])

for (i in c(1:nbs)) {   
    recsynth     <- vapply(bspow_pred, FUN=function(x) mean(rbinom(ns,1,x)), numeric(1))  # synthetic data.
    bsparms[i,]  <- unlist(optim(pout$par,powdiscrep,rec=recsynth,ri=ri)$par)  # estimate values for a, b and c
}

#function to plot a histogram
histoplot<-function(x,l4x) {
    hist(x,xlab=l4x,main="",xlim=c(0,1),cex.lab=1.5,cex.axis=1.5)
    lq <- quantile(x,0.025)
    abline(v=lq,lty="dashed",lwd=2)
    uq <- quantile(x,0.975)
    abline(v=uq,lty="dashed",lwd=2)
    return(c(lq,uq))
}
x11(width=5,height=2)
par(mfcol=c(1,3),las=1)
for (i in c(1:dim(bsparms)[2])) {
    print(histoplot(bsparms[,i],names(sparms)[i]))
}
```


## Model complexity 

```{r}
n <- 1000
m <- seq(1,10,length.out = n)

#pdf(file="polycomplex.pdf", width=7, height=4)

par(mfrow=c(1,2))

for (modOrder in c(2,10)){
  plot(1:10,1:10,xlim=c(0.5,10.5), ylim=c(0,5), type="n",
       xlab="Physical intensity", ylab="Sensation", las=1)
  kk <- poly(x = m,degree = modOrder)
  for (rep in 1:10){
    tcoef <- rnorm(modOrder, sd=5)
    #tcoef <- rexp(modOrder, rate=0.2)
    y <- t(kk%*%tcoef)
    y <- y - min(y)
    lines(x=m, y, type="l", col=rgb(0,0,0,0.5))
  }
}

#dev.off()
```

Out of set prediction error: 

```{r}
n <- 20
n_transfer <- 2
nReps <- 1000
maxOrder <- 6

trainErr <- rep(0, maxOrder)
testErr <- rep(0, maxOrder)

m <- 1:n

tcoef <- c(1,-1,1)
kk <- poly(m,length(tcoef))
realf <- t(kk%*%tcoef)

plot(m, realf)

for (rep in 1:nReps){
  #tcoef <- c(-1,-1,1)
  dat <- realf + rnorm(n, sd=0.5)
  m_t <- sample.int(n, size=n_transfer)
  m <- setdiff(1:n,m_t)
  for (modOrder in 1:maxOrder){
    tTrain <- 0
    tTest <- 0
    lm_t <- lm(dat[m] ~ poly(m,modOrder))
    tTrain <- tTrain + mean(lm_t$residuals^2)
    tTest <- tTest + mean((predict(lm_t, data.frame(m=m_t))-dat[m_t])^2)
    trainErr[modOrder] <- trainErr[modOrder] + tTrain
    testErr[modOrder] <- testErr[modOrder] + tTest
  }
}

kk <- cbind(trainErr,testErr)/nReps

#pdf(file="outOfSet.pdf",width=4,height=5)
matplot(kk, type="l",lty=1,
        ylim=c(0,0.7),
        lwd=2,col=c("black","grey"),
        xlab="Complexity",ylab="Prediction Error")
legend(1,0.7,c("Train","Test"),lwd=2,lty=1,col=c("black","grey"))
#dev.off()
```

## Polynomial law of sensation 

```{r}
n <- 10
stimrange <- c(1,10)
m <- seq(stimrange[1],stimrange[2],length.out = n)
s <- log(m) + rnorm(n,sd=0.5)
s[s < 0] <- 0

lm_3 <- lm(s ~ poly(m,3))
lm_n <- lm(s ~ poly(m,n-1))

#pdf(file="Nihm76.pdf", width=6, height=6)

par(mfrow=c(1,2))
# plot(m,s, pch=3, las=1,
#      xlab='Physical Intensity',
#      ylab='Perceived Intensity')
# points(predict(lm_3))
# 
# plot(m,s, pch=3, las=1,
#      xlab='Physical Intensity',
#      ylab='Perceived Intensity')
# points(predict(lm_n))

m_fill <- seq(stimrange[1],stimrange[2],length.out = 1000)
plot(m,s, pch=3, las=1,
     xlab='Physical Intensity',
     ylab='Perceived Intensity')
lines(m_fill,
      predict(lm_3,data.frame(m=m_fill)),
      type="l", lty=1, las=1)

plot(m,s, pch=3, las=1,
     xlab='Physical Intensity',
     ylab='Perceived Intensity')
lines(m_fill,
      predict(lm_n,data.frame(m=m_fill)),
      type="l", lty=1, las=1)

#dev.off()
```


## Bias-variance 

```{r}
n <- 20
n_transfer <- 10
nReps <- 500
maxOrder <- 10
realOrder <- 3

m <- 1:n

bias <- rep(0, maxOrder)
variab <- rep(0,maxOrder)

tcoef <- c(1,-1,1)
kk <- poly(m,length(tcoef))
realf <- kk%*%tcoef

allPreds <- {}
for (modOrder in 1:maxOrder){
  allPreds[[modOrder]] <- matrix(rep(0,n*nReps),ncol=n)
}

for (rep in 1:nReps){
  #tcoef <- c(-1,-1,1)
  dat <- realf + rnorm(n, sd=0.5)
  for (modOrder in 1:maxOrder){
    lm_t <- lm(dat ~ poly(m,modOrder))
    allPreds[[modOrder]][rep,] <- predict(lm_t)
  }
}

#pdf(file="biasVarianceIllus.pdf",width=7,height=7)
par(mfrow=c(2,2))
for (ii in c(1,3,7,10)){
  plot(m,realf,ylim=c(-1.5,1),
       type="b",lwd=2,las=1,
       xlab="x", ylab="y")
  lines(colMeans(allPreds[[ii]]),
        col=rgb(0,0,0,0.5),lwd=3)
  kk <- apply(allPreds[[ii]],2,function(x) quantile(x,c(0.1,0.9)))
  lines(smooth(kk[1,]), lwd=2,lty=2,col=rgb(0,0,0,0.5))
  lines(smooth(kk[2,]), lwd=2,lty=2,col=rgb(0,0,0,0.5))
  legend(10,-0.5,legend=c("true","fitted"), lty=1, 
         col=c("black","grey"),lwd=2)
  title(paste("Order = ",ii))
}
#dev.off()

bias <- rep(0,maxOrder)
variab <- rep(0,maxOrder)
for (ii in 1:maxOrder){
  mm <- colMeans(allPreds[[ii]])
  bias[ii] <- mean((mm-realf)^2)
  kk <- apply(allPreds[[ii]],1,function(x) (x-mm)^2)
  variab[ii] <- mean(rowMeans(kk))
}

#pdf(file="biasVarTradeOff.pdf",width=5,height=5)
par(mfrow=c(1,1))
plot(bias+variab, type="l", lwd=2, lty=2,
     ylim=c(0,0.15), 
     xlab="Complexity", ylab="Error")
lines(bias,lwd=2)
lines(variab,lwd=2,col=rgb(0,0,0,0.5))
legend(6,0.07,
       legend=c("Total Error",expression("Bias" ^ "2"),"Variance"),
       lty=c(2,1,1),lwd=1,col=c("black","black","grey"))
#dev.off()
```


## Bayesian updating 

### Conjugate normal updating 

```{r}
# Parameters for prior
mu_0 <- 0       # prior mean
sigma_0 <- 0.3    # prior standard deviation

# Parameters for data (likelihood)
sigma <- 0.5               # known standard deviation of data
# x_data <- c(0.4, 0.5, 0.6) # observed data
set.seed(9834)
x_data <- rnorm(7, 0.8, sigma)
n <- length(x_data)
x_bar <- mean(x_data)

# Posterior parameters
sigma_n_sq <- 1 / (1/sigma_0^2 + n/sigma^2)
mu_n <- sigma_n_sq * (mu_0/sigma_0^2 + n * x_bar / sigma^2)
sigma_n <- sqrt(sigma_n_sq)

# Grid for plotting
theta <- seq(-1, 2, length.out = 500)
prior_density <- dnorm(theta, mean = mu_0, sd = sigma_0)
likelihood_density <- dnorm(theta, mean = x_bar, sd = sigma / sqrt(n))  # Likelihood ~ N(x̄, σ²/n)
posterior_density <- dnorm(theta, mean = mu_n, sd = sigma_n)

maxy <- max(c(prior_density, likelihood_density, posterior_density))

# Plot
plot(theta, prior_density, type = "l", col = "blue", lwd = 2,
     ylim = c(0, maxy * 1.1),
     ylab = "Density", xlab = expression(theta),
     main = "Conjugate Normal-Normal Bayesian Updating")

points(x = x_data, y = rep(maxy * 0.1, length(x_data)))

lines(theta, likelihood_density, col = "red", lwd = 2)
lines(theta, posterior_density, col = "darkgreen", lwd = 2)

legend(x = max(theta) * 2/3, y = maxy * 0.8,
       legend = c("Prior", "Likelihood", "Posterior"),
       col = c("blue", "red", "darkgreen"),
       lwd = 2, lty = 1, box.lwd = 1, box.col = NA, bg = NA)
```


### Urn model 

```{r betas}
x <- seq(0, 1, by = 0.01)
stages <- list(c(0.5, 0.5), c(1, 1), c(1, 2), c(5, 3), c(16, 12))

curdat <- lapply(stages, FUN = function(ab){
  data.frame(x = x, 
              stage = paste0(ab, collapse = ","),
                    dens = dbeta(x, ab[1], ab[2]))
})

betdat <- do.call(rbind, curdat)

betdat$stf <- factor(betdat$stage, levels = sapply(stages, paste0, collapse = ","), ordered = TRUE)

for(st in levels(betdat$stf)[-1]){
  

  bet1_2 <- betdat %>%
    filter(stage != "0.5,0.5") %>%
    ggplot(aes(x = x, y = dens, color = stf, alpha = interaction(stf < st, stage == st, drop = TRUE))) +
    geom_line(linewidth = 1) +
    scale_color_brewer(palette = "Dark2" ) +
    scale_alpha_manual(values = c("TRUE.FALSE" = 0.4, "FALSE.FALSE" = 0, "FALSE.TRUE" = 1),
      guide = "none") +
    labs(color = "") +
    theme_classic()

  bet1_2
  
  ggsave(sprintf("_figures/beta%s.png", gsub(",|\\.", "_", st)), height = 12, width = 12, unit = "cm")  

}



```

### Metropolis 

```{r met-elreath}
num_weeks <- 1e5
positions <- rep(0,num_weeks)
current <- 10
for ( i in 1:num_weeks ) {
# record current position
positions[i] <- current
# flip coin to generate proposal
proposal <- current + sample( c(-1,1) , size=1 )
# now make sure he loops around the archipelago
if ( proposal < 1 ) proposal <- 10
if ( proposal > 10 ) proposal <- 1
# move?
prob_move <- proposal/current
current <- ifelse( runif(1) < prob_move , proposal , current )
}

barplot(table(positions))
```

```{r met-balls}
#perform MCMC

# Observed data
n <- 28        # number of trials
k <- 16        # number of red balls observed

# Prior: Beta(alpha, beta)
alpha <- 1
beta <- 1

# Metropolis-Hastings parameters
iterations <- 10000
burnin <- 1000
theta_current <- 0.5  # current parameter.
proposal_sd <- 0.05

# Storage for samples
samples <- numeric(iterations)

# Log posterior function (up to constant)
log_posterior <- function(theta, k, n, alpha, beta) {
  if (theta <= 0 || theta >= 1) return(-Inf)  # Reject invalid theta
  return((k + alpha - 1) * log(theta) + (n - k + beta - 1) * log(1 - theta))
}

# Metropolis-Hastings sampling
set.seed(123)  # For reproducibility
for (i in 1:iterations) {
  # Propose new theta from normal distribution
  theta_proposal <- rnorm(1, mean = theta_current, sd = proposal_sd)
  
  # Calculate acceptance ratio (on log scale)
  log_r <- log_posterior(theta_proposal, k, n, alpha, beta) -
           log_posterior(theta_current, k, n, alpha, beta)
  
  # Accept or reject
  if (log(runif(1)) < log_r) {
    theta_current <- theta_proposal
  }
  
  samples[i] <- theta_current
}

# Discard burn-in
samples_post <- samples[(burn_in + 1):iterations]

# Plot posterior samples:

par(mfrow = c(1, 2))

plot(samples_post, type = "l")

hist(samples_post, breaks = 50, probability = TRUE, col = "skyblue",
     main = "Posterior of θ via Metropolis-Hastings", xlab = expression(theta))
abline(v = mean(samples_post), col = "red", lwd = 2)

par(mfrow = c(1, 1))


```


```{r}
# Example for assignment:
x <- rnorm(100, 3, 4)
y <- 0 + rnorm(100, x, 4) * 0.3


plot(x, y)

iter <- 1e5
positions <- rep(0, iter)  # parameter estimates.
current <- 0

for ( i in 1:num_weeks ) {
  # record current position
  positions[i] <- current
  # flip coin to generate proposal
  proposal <- current + rnorm(1, 0, 1)  # proposal for parameter.

  # Likelihood (no prior for now): 
    


  # now make sure he loops around the archipelago
  if ( proposal < 1 ) proposal <- 10
  if ( proposal > 10 ) proposal <- 1
  # move?
  prob_move <- proposal/current
  current <- ifelse( runif(1) < prob_move , proposal , current )
}

barplot(table(positions))
```


