---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 2: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for two.

1. Upload your assignments to the joint git repository. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


# Task 1: Simulating data 

1. Pick an environment: gambles, probability learning [**OR?**; **AND: REALLY DO THIS OR PICK ONE?**]

1. Pick a model **CPT...** 

1. Visualize the function(s) you use 

1. Simulate data from a few hundred individuals with different parameters (if)


## Option 1: Gamble data 

These data represent choices between gambles.
According to a model, individuals choose one gamble over the other.

Each gamble $j$ consists of a probability $p_j$ and an outcome $o_j$ (for sake of simplicity the alternative outcome is zero and does not affect the calculated value).


```{r generate-gambles}
ngamble <- 50

probs <- round(runif(ngamble * 3, 0, 1), 2)  # Sample probabilities from a uniform distribution.
outcomes_pos <- round(runif(ngamble, 0, 100))
outcomes_neg <- round(runif(ngamble, -100, -1))
outcomes_mix <- round(runif(ngamble, -100, 100))

# Create gamble pairs to simulate choices:
gpairs <- data.frame(p = matrix(probs, ncol = 2),
                     o = rbind(matrix(outcomes_pos, ncol = 2),
                               matrix(outcomes_neg, ncol = 2),
                               matrix(outcomes_mix, ncol = 2)),
                     type = rep(c("gain", "loss", "mix"), each = ngamble/2)
                     )
```


Pick one model *in addition* to Expected Value (EV), which has no parameters.

**[TODO: PROVIDE GAMBLES IN THE LOASS AN GAIN DOMAIN BUT ALLOW TO FILTER -- should I select random gambles or interesting gambles?]**


### Model suggestions  

All with $p_i$ being the probability of (monetary) outcome $x_i$ and $k$ being the number of outcomes in the gamble.

#### Expected value (EV) 

\[
EV = \sum_{i=1}^k p_i x_i
\]


#### Expected utility (EU) 

\[
EV = \sum_{i=1}^k p_i x_i^{\alpha}
\]

with $\alpha < 1$ representing diminishing marginal utility (additional money having less value than the previous dollar).



#### Cumulative prospect theory 

The subjective value $V$ of an option consists of the sum of decisions weights $\p_i$ times outcomes passed through a function $v(x_i)$ with $v(0) = 0$ [e.g., @GloecknerPachur12]

\[
V = \sum_{i=1}^k \pi_i v(x_i)
\]

Functions differ for positive and negative outcomes.
You can decide if you want to focus on the gain-domain (positive outcomes) by filtering the inputs ("gambles").


I suggest the following functions (You can look for your own functions in the literature): 

Probability weighting function:

Probabilities are weighted by:

 \[\pi(p_i) = \frac{p_i^c}{[p_i^c + (1 - p_i^c)]^{1/c}}\] 
 
with $c = \gamma$ for gains and $c = \delta$ for losses. For simplicity I suggest to assume that $\gamma = \delta < 1$.


Value functions:

\[
\begin{array}{rcl}
v(x) = x^{\alpha} & \mbox{for} & x \geq 0 \\ 
v(x) = -\lambda(-x)^\beta & \mbox{for} & x < 0 \\
\end{array}
\]

For simplicity I suggest to assume that $\alpha = \beta < 1$ and maybe even $\lambda = 1$ (i.e., no stronger weighting of losses resulting in *loss aversion*).



#### Simulating choices 

If you want to simulate choices you can use a softmax rule: 

\[p(A) = \frac{e^{\phi v(A)}}{e^{\phi v(A)} + e^{\phi v(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 



### [YOUR ASSIGNMENT HERE]



## Option 2: Probability learning data 

In each round one of two options is rewarded with a unit reward (i.e., one point) the other is unrewarded.
Individuals want to maximize their rewards.

```{r generate-pdat}
probdat <- seq(0.05, 0.95, by = 0.15)  # probabilities.

# FUnction to generate rewards:
curprob <- probdat[1]  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.
```



### Model suggestions 


#### Probability matching 

Individuals select the options relative to their probability of being rewarded. 



#### Simple reinforcement learning ($Q$-learning)

Individuals *learn* about the best option. In each round $t$ they update a representation $Q$ of each option (also called "action").

\[
Q(t+1) = Q(t) + \alpha (r_t - Q_t)
\]

This means that the expectation at time $t+1$ is the previous expectation plus the weighted "surprise" about the outcome (reflected in the deviation of the outcome from the expectation).
The impact of this is governed by the learning rate $\alpha \in [0;1]$.


#### Simulating choices 

If you want to simulate choices from $Q$-values for options A and B, you can use a softmax rule: 

\[p_t(A) = \frac{e^{\phi Q_t(A)}}{e^{\phi Q_t(A)} + e^{\phi Q_t(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 


### [YOUR ASSIGNMENT HERE]

Simulate choices from your model of choice for different probabilities from `probdat`.


# [TODO] Task 2: Grid search 

**MAYBE also! REQUIRES SOME CHOICE DATA etc. -- use simulated data from week before**




# Task 3: Learn about Bayesian Statistics

Watch the first 9 videos of the following playlist (it is okay, if you don't watch all unti next session):
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6


# Task 4 (optional) 

Pick another model and or environment and simulate data from this. 