---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 2: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for one ungraded assignment.

1. Upload your assignments to ilias until Wednesday 3 pm. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 






# Task 1: Getting data from the previous assignment

Take the model you used in the previous assignment and 

1. repeat the visualizations of the function(s) involved (i.e., the relationships between values and probabilities to the values used), 
2. visualize the data and the choices you simulated for each of the gambles (the head of a table is sufficient).

You may use a subset of the parameters and gambles you simulated data for.
Use at least two parameter combinations.

Use this opportunity also to fix any errors (especially those in my feedback). 

*You may revert to a simpler model (i.e., from prospect theory to expected utility).*



# Task 2: Parameter estimation

Use your simulated choice data to estimate the parameters of your models by means of maximum likelihood estimation.

Hint: You get a predicted choice probability based on a combination of parameters as model output (through your softmax rule).
You can use this probability to get the likelihood of a given choice through a bernoulli distribution `dbinom(data, prob)` -- how likely is it to observe the choice, given the probability output by the model. 
You then try to find those paramter combination that minimizes the negative *log* likelihood (or rather twice the log likelihood).

```{r obtaining-minimzing-likelihood}
# Obtaining and minimizng likellihood:
p <- 0.4  # probability (model prediction).

vchoice <- c(0, 1, 0, 0, 0, 1)  # vector of choices.

# Likelihood of choice 1:
  (lik <- dbinom(vchoice[1], 1, p))  # show likelihood of one choice.

# Likelihood of all choices:
  (lik <- dbinom(vchoice, 1, p)) 

# To maximize, we use the log-likelihood,
  # which you can obtain like this:
    (ll <- dbinom(vchoice, 1, p, log = TRUE))
  # or this:  
    (ll <- log(dbinom(vchoice, 1, p)))
  # Since minimzation is easier, we minimize 2 timese the negative likelihood (also called deviance):

    (dev <- -2 * sum(ll))  # this is your discrepancy function.
      # Note: Summing up the log likelihoods is equal to multiplying the likelihoods; the latter may result in underflow.
  
```


*In case minimizing the deviance does not work, you may use the choice probabilities directly (and also a RMSD).*


## Use a grid search 



## Use the Nelder-Mead Simpex algorithm


# Task 3: Learn about Bayesian Statistics

Watch the first 9 videos of the following playlist (it is okay, if you don't watch all unti next session):
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6

