---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 2: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for one ungraded assignment.

1. Upload your assignments to ilias until Wednesday 3 pm. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 



## Probability learning data 

In each round one of two options is rewarded with a unit reward (i.e., one point) the other is unrewarded.
Individuals want to maximize their rewards.

```{r generate-pdat}
probdat <- seq(0.05, 0.95, by = 0.15)  # probabilities.

# FUnction to generate rewards:
curprob <- probdat[1]  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.
```

### Model suggestions 


#### Probability matching 

Individuals select the options relative to their probability of being rewarded. 



#### Simple reinforcement learning ($Q$-learning)

Individuals *learn* about the best option. In each round $t$ they update a representation $Q$ of each option (also called "action").

\[
Q(t+1) = Q(t) + \alpha (r_t - Q_t)
\]

This means that the expectation at time $t+1$ is the previous expectation plus the weighted "surprise" about the outcome (reflected in the deviation of the outcome from the expectation).
The impact of this is governed by the learning rate $\alpha \in [0;1]$.


#### Simulating choices 

If you want to simulate choices from $Q$-values for options A and B, you can use a softmax rule: 

\[p_t(A) = \frac{e^{\phi Q_t(A)}}{e^{\phi Q_t(A)} + e^{\phi Q_t(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 





### [YOUR ASSIGNMENT HERE]

Simulate choices from the model for different probabilities from `probdat`.


# [TODO] Task 1: Simulate your data  

1. Simulate data from a probability learning environment 

1. Pick a model, e.g., simple reinforcement learning 

1. Visualize the function(s) you use 

1. Simulate data from a few hundred individuals with different parameters (if)



# [TODO] Task 2: Grid search 




# Task 3: Learn about Bayesian Statistics

Watch the first 9 videos of the following playlist (it is okay, if you don't watch all unti next session):
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6

