---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 4: Simulation, Estimation and Paramter Uncertainty'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for one ungraded assignment.

1. Upload your assignments to ilias until Wednesday 3 pm. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


## Probability learning data 

In each round one of two options is rewarded with a unit reward (i.e., one point) the other is unrewarded.
Individuals want to maximize their rewards.

```{r generate-pdat}
# Function to generate rewards:
curprob <- 0.05  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.
```


### Model: Individual Reinforcement Learning 

Individuals *learn* about the best option $i$. In each round $t$ they update a representation $Q$ of each option (also called "action").

\[
Q_{t+1}(i) = Q_t(i) + \alpha [r_t - Q_t(i)]
\]

This means that the expectation at time $t+1$ is the previous expectation plus the weighted "surprise" about the outcome (reflected in the deviation of the outcome from the expectation).
The impact of this is governed by the learning rate $\alpha \in [0;1]$.


#### Simulating choices 

If you want to simulate choices from $Q$-values for options A and B, you can use a softmax rule: 

\[p_t(A) = \frac{e^{\phi Q_t(A)}}{e^{\phi Q_t(A)} + e^{\phi Q_t(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 





# Task 1: Simulate your data  

* simulate 100 choices for at least 10 individuals with different parameter combinations in a probability learning environment, where on each trial one option is rewarded with a probability $p$ and the other with a probability $p-1$ (choose a reward probability for the options, wit $p \notin {0;0.5,1}$)

* Simulate these choices from a simple reinforcement learning model

    + For each individual simulate its own rewards from a probability learning environment -- take care of temporal dependencies (a non-chosen option cannot be rewarded); assume that the rewarded option provides a unit reward of 1 

    + vary each parameter so that you get at least 2x2 combinations; for each parameter combination  
    
* visualize the simulated choice data over time for at least 3 individuals with at least two different paramter combinations (and note down their respective parameters) 




# Task 2: Parameter recovery

* Use model fitting on your simulated choice data to recover the parameters $\alpha$ and $\phi$ you used *for each individual* 

* Use the maximum likelihood approach and one algorithm *one*: a grid search or the Nelder-Mead algorithm 

* Visualize the relationship between the best fitting parameters with the true parameters for each individual 



# Task 3: Parameter uncertainty 

* For each of the two parameters $\alpha$ and $\phi$ in the previous model estimate the variability (for instance with bootstrapping) for at least 3 individuals.

* Visualize the parameter variability (e.g., as histograms or violin plots) 




# Task 4: Learn about Bayesian Statistics

Make sure to have watched until video 9 of the following playlist:
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6

