---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 2: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for one ungraded assignment.

1. Upload your assignments to ilias until Wednesday 3 pm. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 


## Probability learning data 

In each round one of two options is rewarded with a unit reward (i.e., one point) the other is unrewarded.
Individuals want to maximize their rewards.

```{r generate-pdat}
probdat <- seq(0.05, 0.95, by = 0.15)  # probabilities.

# FUnction to generate rewards:
curprob <- probdat[1]  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.
```

### Model: Individual Reinforcement Learning 

Individuals *learn* about the best option. In each round $t$ they update a representation $Q$ of each option (also called "action").

\[
Q(t+1) = Q(t) + \alpha (r_t - Q_t)
\]

This means that the expectation at time $t+1$ is the previous expectation plus the weighted "surprise" about the outcome (reflected in the deviation of the outcome from the expectation).
The impact of this is governed by the learning rate $\alpha \in [0;1]$.


#### Simulating choices 

If you want to simulate choices from $Q$-values for options A and B, you can use a softmax rule: 

\[p_t(A) = \frac{e^{\phi Q_t(A)}}{e^{\phi Q_t(A)} + e^{\phi Q_t(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 





# Task 1: Model comparison 

Take your simulated choice data from the last assignment 

* Determine how well a model according to which individuals select each option with a fixed probability describes your data.

* Determine the best fitting probability for each individual.

* Compare the fit of this simple model to the fit of your learning model from the last assignment. What do you observe?


# Task 2: Win-stay, lose shift

Determine how well a simple win-stay, lose-shift (WSLS) model describes your data.
According to WSLS, individuals switch, when they receive no reward and select the same option when they were rewarded.

* What is the problem with such a model?

* Determine the fit of the WSLS model to your simulated choice data. What do you need to do (hint: you need some probability)?

* Compare the fit of the WSLS model to your learning model. What do you observe?

*WSLS does not need a free parameter. However, you may estimate up to two probabilities: one after "win" (being rewarded) and one after "lose" (not being rewarded.*


# Task 6: Learn about Bayesian Statistics

Make sure to have watched until video 17 of the following playlist:
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6

