---
title: 'Computational modelling of judgments, decisions and cognition -- Assignments 2: Getting started'
author: "[YOUR NAME]"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r markdown-prep}
library(dplyr)  # basic package for changing data structures.

library(ggplot2)  # important package for plotting.
```


# General rules for handing in assignments 


1. You have to hand in all assignments except for one ungraded assignment.

1. Upload your assignments to ilias until Wednesday 3 pm. 

1. Assignments are *not* graded, unless it says so in the beginning of the assignment. 

1. You have to complete all numbered tasks; if there are one or more bonus tasks, you may skip them 

**This assignment will be graded** 

## Probability learning data 

In each round one of two options is rewarded with a unit reward (i.e., one point) the other is unrewarded.
Individuals want to maximize their rewards.

```{r generate-pdat}
probdat <- seq(0.05, 0.95, by = 0.15)  # probabilities.

# FUnction to generate rewards:
curprob <- probdat[1]  # selecting a probability.

tr <- 200  # 200 rounds.
rbinom(tr, 1, curprob)  # get an outcome according to a bernoulli distribution.
```

### Model: Individual Reinforcement Learning 

Individuals *learn* about the best option. In each round $t$ they update a representation $Q$ of each option (also called "action").

\[
Q(t+1) = Q(t) + \alpha (r_t - Q_t)
\]

This means that the expectation at time $t+1$ is the previous expectation plus the weighted "surprise" about the outcome (reflected in the deviation of the outcome from the expectation).
The impact of this is governed by the learning rate $\alpha \in [0;1]$.


#### Simulating choices 

If you want to simulate choices from $Q$-values for options A and B, you can use a softmax rule: 

\[p_t(A) = \frac{e^{\phi Q_t(A)}}{e^{\phi Q_t(A)} + e^{\phi Q_t(B)}}\]

with $\phi$ indicating how noisy the responding is, with smaller $\phi$ being related to more random responding (inverse temperature parameter). 


# Task 1: Revisit the previous assignment 

1. Recover the parameters, thereby try to: 
    + use an algorithm like Nelder-mead simplex 
    + calculate the $-2LL$ (deviance) based on a vector of choice probabilities (hint: you can create a vector of choices for each round; you can then use this vector to index the choice probabilities)
    + what do the parameters mean? 
    
```{r example-mat}
set.seed(94385)
choices <- rbinom(5, 1, 0.7) 

p0 <- round(rbeta(5, 9, 9), 2)  # random probabilities for option 0 (probability in column 1). 
pmat <- cbind(p0, 1-p0)  # matrix of choice probaiblities. 

# Obtain the choice probabilities of the chosen option for each trial:
p_chosen <- pmat[cbind(1:length(choices), choices+1)]  # use the choice matrix as index for the probability matrix (note: in R indexing starts at 1).

# Here, you can also substract the maximum probability in each row, to avpoid overflow.

# To obtain the likelihood of these choices:
-sum(log(p_chosen))
```

2. Redo the bootstrapping (for 3 individuals; ideally with different parameter combinations): This time make sure to 
    a. simulate from the best fitting parameters for the individual (reward information may remain the same) 
    b. refit the model 
    c. obtain the bootstrap estimates
    
Hint: Start with one individual and scale up; also start with few iterations and eventually scale up to about 1,000 

Also make sure to include the best-fitting and true parameter estimates in any plots of the bootstrap estiamtes.



# Task 2: Model comparison 

Take your simulated choice data from the last assignment 

* Determine how well a model according to which individuals select each option with a fixed probability describes your data.

* Determine the best fitting probability for each individual.

* Compare the fit of this simple model to the fit of your learning model from the last assignment. What do you observe?


# Task 3: Win-stay, lose shift (ungraded)

Determine how well a simple win-stay, lose-shift (WSLS) model describes your data.
According to WSLS, individuals switch, when they receive no reward and select the same option when they were rewarded.

* What is the problem with such a model?

* Determine the fit of the WSLS model to your simulated choice data. What do you need to do (hint: you need some probability)?

* Compare the fit of the WSLS model to your learning model. What do you observe?

*WSLS does not need a free parameter. However, you may estimate up to two probabilities: one after "win" (being rewarded) and one after "lose" (not being rewarded.*


# Task 4: Learn about Bayesian Statistics

Make sure to have watched until video 17 of the following playlist:
https://youtube.com/playlist?list=PLFDbGp5YzjqXQ4oE4w9GVWdiokWB9gEpm&si=emFyWAjPCWcwEFs6

